Perhaps the most elegant model of intuitionistic linear logic is
Benton's linear/non-linear (LNL) models \cite{Benton:1994}. On the
semantic side, LNL models are very compact and intuitive, but on the
syntactic side, they correspond to a logic and type theory, called LNL
logic, that allows one to reason with or without weakening and
contraction without the need to annotate every formula of the logic
with the of-course!  modality.  This is possible because formulas like
$A \to B$ are primitives of the logic, and not encoded using the
of-course! modality.  Therefore, we ask the question, ``can a similar
elegant, intuitive, and flexible model and corresponding logic be
defined for the Lambek Calculus?''

LNL models are symmetric monoidal adjunctions, $\cat{C} : F \dashv G :
\cat{I}$, where $\cat{C}$ is a symmetric monoidal closed category with
weakening and contraction, hence is a cartesian closed category, and
$\cat{I}$ is a symmetric monoidal closed category.  Thus, a LNL model
consists of a model of intuitionistic logic on the left and a model of
intuitionistic linear logic on the right related via a pair of adjoint
functors.  Then Benton shows that the of-course! modality can be
recovered by $!A = F(G\,A)$.  These models suggest a more general
framework for working with structural rules.

Our main contributions in this paper are to the Lambek Calculus, but
we make use of a new more general framework that will put this work in
line with planned future work.  Suppose $\cat{M}_1$ and $\cat{M}_2$
are two categories with a bifunctor $\odot_i : \cat{M}_i \times
\cat{M}_i \mto \cat{M}_i$, a distinguished object $I_i \in
\mathsf{Obj}(\cat{M}_i)$, and the two natural isomorphisms $\lambda_A
: A \odot I_i \mto A$ and $\rho_A : I_i \odot A \mto A$.  We call
these categories \emph{magmoidal categories with a unit}.  Examples of
magmoidal categories with a unit are monoidal categories, symmetric
monoidal categories, and cartesian closed categories.

Benton's LNL models can be generalized to magmodial categories.
Simply, take an adjunction $\cat{M}_1 : F \dashv G : \cat{M}_2$ called
an \emph{adjoint model} where the functors $F$ and $G$ preserve the
magmoidal structure similarly to monoidal functors, but without the
coherence diagram for the associator.  Now if we add to $\cat{M}_1$
and $\cat{M}_2$ some structural rules, for example, by making
$\cat{M}_1$ a symmetric monoidal category and $\cat{M}_2$ a monoidal
category, then we obtain an adjoint model that corresponds to a logic
where intuitionistic linear logic ($\cat{M}_1$) and the Lambek
calculus ($\cat{M}_2$) are mixed similarly to LNL models.  However,
this also leads to a new solution to an existing problem.

Linguists \cite{?} have wondered if a modality for exchange can be
added to the Lambek Calculus similarly to how the of-course! modality
adds weakening and contraction to linear logic.  de Paiva and
Eades~\cite{?} show that this is possible by adding a modality $\kappa
A$ with the property that $(\kappa A \otimes B) \multimap (B \otimes
\kappa A)$.  The adjoint model introduced above induces the modality
$eA = F(G\,A)$ with the property $eA \otimes eB \multimap eB \otimes
eA$.  Thus providing a second solution to the problem.

Linguists as well as computer scientists have found many applications
of substructural logics that contain more than one tensor product
satisfying different structural rules.  For example, a commutative
tensor product and a non-commutative tensor product within the same
logic.  An example of such a logic is the logic of bunched
implications, but there sequents $\Gamma \vdash A$ depend on contexts
$\Gamma$ that are trees with multiple different branching nodes
instead of lists.  The branching nodes are necessary to accommodate
multiple different tensor products with different structural
properties.  We introduce, but leave the details to future work, the
idea of using adjoint models to build similar logics, but without the
need for multiple different branching nodes.  Furthermore, adjoints
compose, and this property can be exploited to compose modalities.

We introduce the idea above of having a modality for exchange, but
what about associativity, weakening and contraction?  Indeed it is
possible to give modalities for these structural rules as well using
adjoint models.  Now that we have each structural rule isolated into
their own modality is it possible to put them together to form new
modalities that combine structural rules?  The answer to this question
has already been show to be positive, at least for weakening and
contraction, but we extend this line of work to include exchange, and
in the future associativity as well.  Jacobs~\cite{JACOBS199473} used
monads and distributive laws to relate modalities, but
Melli{\'e}s~\cite{Mellies:2004} shows how to use adjunctions to
accomplish the same, but in a more intuitive and natural way.  Thus,
this part of our work is a natural extension of Melli{\'e}s'.

This work can be used to define a general core logic that is capable
of encoding several different substructural logics.  Some examples are
intuitionistic linear logic, affine logic, contraction logic, and the
(non-associative) Lambek Calculus.  This does have some similarities
to Licata et al.~\cite{licata2017fibrational}, but there they embrace
the tree based structures with multiple different branching node for
contexts from the logic of bunched implications.  However, our work is
trying to understand if this is truly necessary, and if adjunctions
provide a means of, at the very least, hiding these multiple branching
nodes.  Thus, adjunctions may lead to a simplification.

%%% Local Variables: 
%%% mode: latex
%%% TeX-master: main.tex
%%% End:
